The hydra and the rooster

### 2020-07-07 latex/chapter-skippable.tex
work on type classes for ordinal notations

Finally, it is interesting to show that the segment $[0,n[$ is a ``sub-segment'' of
$[0,n+1[$. Since types $t\;n$ are $t\;(n+1)$ are not convertible, we consider a ``cast'' 
function from $t\;n$ to $t\;(n+1)$, and prove this function is compatible with
comparison function of both types, and that it is a bijection from $t\;n$ to
the segment $[0,n-1[$ of $t\;(n+1)$.

In other words, \texttt{FinOrd $n$}  is a subnotation of \texttt{FinOrd $(n+1)$}.

\begin{Coqsrc}
Class  SubNotation {A:Type}
       {ltA: relation A}{stoA:StrictOrder ltA}
       {compareA : A -> A -> comparison}
       (notA : OrdinalNotation stoA compareA)
       {B:Type}{ltB: relation B}{stoB:StrictOrder ltB}
       {compareB : B -> B -> comparison}
       (notB : OrdinalNotation stoB compareB)
       (bound :  B)
       (cast : A -> B):=
  {
  commutation :forall a a' : A,  compareB (cast a) (cast a') =
                                 compareA a a';
  incl : forall a, ltB (cast a) bound;
  surj : forall b, ltB b bound -> exists a:A, cast a = b}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Program Definition cast {i j : nat}  (H: i < j) (alpha: t i) : t j :=
  alpha.

Lemma cast_compare_commute (i j  :nat)(H : i < j) :
  forall alpha beta, compare alpha beta = compare (cast  H alpha)
                                                  (cast  H beta).

@@

We are now able to build an instance of \texttt{SubNotation}

\begin{Coqsrc}
(** biggest element of t (S n) *)

Program Definition biggest_S (n:nat) : t (S n) :=n.
Next Obligation. 
  red;rewrite  Nat.ltb_lt; auto with arith.
Defined.

Instance F_incl n : SubNotation  (FinOrd n) (FinOrd (S n)) (biggest_S n)
                                 (cast (Nat.lt_succ_diag_r n)).
\end{Coqsrc}

\begin{todo}
Comments on \texttt{cast, compare, bigO}

Show that there is no interesting arihmetic, due to the lack of total succesor, addition, etc.
\end{todo}
\begin{itemize}
\item $i\;(i\in\mathbb{N})$ \coq's type \verb@{j:nat | j < i}@ (see also
\href{https://www.math.nagoya-u.ac.jp/~garrigue/lecture/2018_AW/mathcomp-1.7.0/htmldoc/mathcomp.ssreflect.fintype.html#Ordinal}{finite ordinals in Mathcomp}).
\item $\omega$: \coq's type \texttt{nat}, with the order \texttt{Nat.lt}.
\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap/T1}).
\item $\Gamma_0$: The set of terms in Veblen normal form.
\end{itemize}

>>>>

Finally, it is interesting to show that the segment $[0,n[$ is a ``sub-segment'' of
$[0,n+1[$. Since types $t\;n$ are $t\;(n+1)$ are not convertible, we consider a ``cast'' 
function $\iota$ from $t\;n$ to $t\;(n+1)$, and prove this function is  a bijection  from $t\;n$ to
the segment $[0,n-1[$ of $t\;(n+1)$.

This notion is defined generically as a class in
 Module~\href{../src/html/hydras.Prelude.Ordinal-generic.html}{Prelude.Ordinal\_generic} (see Fig~\vref{figure:subsegment}).

\begin{Coqsrc}
(** The segment associated with nA is isomorphic to
    the interval [0,b[ *)

Class  Subsegment 
       `(nA : @OrdinalNotation A ltA stoA compareA)
       `(nB : @OrdinalNotation B ltB stoB compareB)
       (b :  B)
       (iota : A -> B):=
  {
  subseg_compare :forall x y : A,  compareB (iota x) (iota y) =
                                 compareA x y;
  subseg_incl : forall x, ltB (iota x) b;
  subseg_onto : forall y, ltB y b  -> exists x:A, iota x = y}.
\end{Coqsrc}

 \begin{figure}[h]
   \centering
   \begin{tikzpicture}[very thick, scale=0.6]
\begin{scope}[color=blue]
\node (A) at (0,0) {$A$};
\node(A0) at (2,0)[label=below:$0$]{$\bullet$};
\node(A1) at (3,0)[label=below:$1$]{$\bullet$};
\node(A2) at (4,0)[label=below:$2$]{$\bullet$};
\node (Adots) at (6,0) {$\ldots$};
\end{scope}
\begin{scope}[color=red]
\node (B) at (0,2) {$B$};
\node(B0) at (2,2)[label=above:$0$]{$\bullet$};
\node(B1) at (3,2)[label=above:$1$]{$\bullet$};
\node(B2) at (4,2)[label=above:$2$]{$\bullet$};
\node (Bdots) at (6,2) {$\ldots$};
\node (b) at (8,2) [label=above:$b$]{$\bullet$};
\node (bsucc) at (9,2) [label=above:$b+1$]{$\bullet$};
\node (Bdots2) at (10,2) {$\ldots$};
\end{scope}
\begin{scope}[color=red!50!blue]
\draw [->,thin] (A0) -- node [auto] {$\iota$} (B0);
\draw [->,thin] (A1) -- node [auto] {$\iota$} (B1);
\draw [->,thin] (A2) -- node [auto] {$\iota$} (B2);
\draw [->,thin] (Adots) -- node [auto] {$\iota$} (Bdots);
\end{scope}
\end{tikzpicture}
   \caption{\textcolor{blue}{$A$} is a sub-segment  of \textcolor{red}{$B$}}
   \label{fig:subsegment}
 \end{figure}

@@

\begin{Coqsrc}
Program Definition iota {i j : nat}  (H: i < j) (alpha: t i) : t j :=
  alpha.

Lemma iota_compare_commute (i j  :nat)(H : i < j) :
  forall alpha beta, compare alpha beta = compare (iota  H alpha)
                                                  (iota  H beta).
\end{Coqsrc}

We are now able to build an instance of \texttt{Subsegment}

\begin{Coqsrc}
Section Inclusion_ij.

  Variables i j : nat.
  Hypothesis Hij : (i < j)%nat.

   Program Definition iota  (alpha: t i) : t j :=  alpha.

   Let b : t j := exist _ i Ltb_ij.

  Instance F_incl_ij  : SubSegment  (FinOrd i) (FinOrd j) b iota.

  End Inclusion_ij.

@@

\begin{remark}
 There is no interesting arihmetic on finite ordinals, since functions like successor, addition;, etc.,  cannot represented in \coq{} as \emph{total} functions.
\end{remark}

\begin{itemize}
\item $i\;(i\in\mathbb{N})$ \coq's type \verb@{j:nat | j < i}@ (see also
\href{https://www.math.nagoya-u.ac.jp/~garrigue/lecture/2018_AW/mathcomp-1.7.0/htmldoc/mathcomp.ssreflect.fintype.html#Ordinal}{finite ordinals in Mathcomp}).
\item $\omega$: \coq's type \texttt{nat}, with the order \texttt{Nat.lt}.
\item $\omega^n$ (for some  integer  $n\geq 2$) : the type of $n$-uples of natural numbers, with the lexicographic product of $n$ copies of \texttt{Nat.lt}.
\item  $\omega^\omega$: the set of nonincreasing sequences of natural numbers, lexicographically ordered (also the set of finite multisets of natural numbers).
\item $\epsilon_0$: The set of terms in Cantor normal form (see Chap.~\ref{chap/T1}).
\item $\Gamma_0$: The set of terms in Veblen normal form.
\end{itemize}
